### 2014-08-14 cover.tex
Add cover letter

\documentclass{article}

\usepackage{url}

\title{Cover Letter to JMLR Machine Learning Open Source Software Submission:\\
BayesPy: Variational Bayesian Inference in Python}

\author{Jaakko Luttinen\\
  \url{jaakko.luttinen@aalto.fi}\\
    Aalto University, Finland}

\begin{document}

\maketitle

\begin{itemize}

\item There are no previous publications by the author that significantly
  overlap with this submission.

\item There are no co-authors to be aware of the current submission and consent
  to its review by JMLR.

\item Possible conflict of interest with the following persons: Antti Honkela,
  Samuel Kaski

\item Suggestions of action editors: Geoffrey Holmes, Cheng Soon Ong, Mark Reid

\end{itemize}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-master: t
%%% End: