### 2015-03-10 response.tex
Add response skeleton

\documentclass{article}

\usepackage{url}

\title{Review Reply to JMLR Machine Learning Open Source Software Submission (14-324):\\
BayesPy: Variational Bayesian Inference in Python}

\author{Jaakko Luttinen\\
  \url{jaakko.luttinen@aalto.fi}\\
    Aalto University, Finland}

\begin{document}

\maketitle

\section{Responses to the Reviewers of Submission 14-324 (v1)}

I would like to thank..

added features etc

% \subsection{Decision (meta-review)}

% The blah blah

% \begin{itemize}

% \end{itemize}

\subsection{Reviewer 1}

\begin{itemize}

\item \underline{syntax}\\
  yeah yeah

\item \underline{speed comparison}\\
  ooh yeeeeah

\item \underline{supported models}\\

\item \underline{scheduling}\\

\end{itemize}

\subsection{Reviewer 2}

\begin{itemize}

\item \underline{installation instructions}\\

\item \underline{user community}\\

\item \underline{speed comparison}

\end{itemize}

\subsection{Reviewer 3}

Reviewer 3 did not point out any BLAH BLAH BLAH

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-master: t
%%% End: